---
title: "Mycobiont data QC"
author: "Neha Sahu, adopted by Gulnara Tagirdzhanova"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Rationale:** this is basic QC for all samples of pure culture mycobiont and lichen. We will use to select which sample we canuse and comparisons

```{r, message=F, include=FALSE}
library(here)
lapply(c("tidyverse","corrplot","dplyr","gplots","limma"), library, character.only = T)
cl<-c("blue","white","red") #this is the color scheme set in the order low, mid, high values
col2<-colorRampPalette(cl)
```

#### Create tpm table from kallisto output
```{r, message=F}
counts<-read.delim2("../analysis_and_temp_files/06_meta_mapping/kallisto_mycobiont_only.txt",sep=" ")
metadata<-read.csv2("../analysis_and_temp_files/06_meta_mapping/metadata_shared_with_neha.csv",sep=",")
length(unique(counts$sample)) #this should be 31 as in the metadat.csv file
counts$tpm<-round(as.numeric(counts$tpm),digits = 3) #round off tpm upto 3 decimals
counts_tab<-pivot_wider(counts[c(1,5,6)],names_from=sample,values_from = tpm) #widen the table
```

## 1. All Samples (_Xanthoria_ and Lichen)
#### Calculate and plot pearson's correlation coefficients between all the samples
```{r fig.width = 10, fig.height = 10, message=F}
all_counts<- counts_tab[rowSums(counts_tab[c(2:32)])>0,] #remove rows which have 0 in all samples
all<-cor(all_counts[c(2:32)], method="pearson",use="pairwise.complete.obs")
ord_all<-c("XBA1","XBA2","XSA1","XSA2","XSA2_2","XTA1","XTA2","XBC1","XBC2","XMC2","XSC1",
       "XSC2","XTC2","XBE1","XBE2","XSE2","XTE2","MP_I","MP_II","KS48XB1","KS48XB2","KS48XB3",
       "KS9XB1","KS9XB2","KS9XB3","S_21XB1","KS21XB1","S_21XB3","S_42XB1","S_42XB2","S_42XB3") #arrange order according to biological replicates (optional)
all1 <- all[, ord_all]
all1<-all1[ord_all, ]

# For correlation plot
corrplot(all1,method="color",type="full",is.corr=F, col=col2(200),addgrid.col = 0,tl.cex = 1, tl.col = "black")

# for PCA
plotMDS(all_counts[c(2:32)],main="All samples") 

# For heatmap
heat1<-column_to_rownames(all_counts, var = "target_id")
exp_matrix<-data.matrix(heat1)
mycols = colorRampPalette(c("blue","white","red"))(100) 
hc_rows <- hclust(as.dist(1-cor(t(exp_matrix), method="pearson")), method="average") # for clustering genes
hc_cols <- hclust(as.dist(1 - cor(exp_matrix, method = "pearson")), method = "average") # to arrange the sample names at x-axis according to hierarchical clustering
htmap<-heatmap.2(exp_matrix, Colv=as.dendrogram(hc_cols),Rowv=as.dendrogram(hc_rows),dendrogram = "both",trace = "none", 
                 col = mycols, keysize = 1, key.title = "Title",
                 scale = c("row"), cexRow = 0.75, cexCol = 1, srtCol = NULL,  
                 srtRow = NULL, adjRow = c(0,NA), sepcolor="white" ,margins = c(7,15), 
                 lwid = c(2,8),
                 lhei= c(2,8), main ="All samples", na.color="black") #this is the final one for the heat map
```

#### Conclusions:
* Lichen and culture samples broadly form two groups, but with two big buts:
* 2dpi cultures differ dramatically from the rest
* MP_I and MP_II lichen samples ("filter") group with the cultures (in the cloud of >2dpi). Perhaps, they were incubated for too long in the lab and weren't happy?


## 2. Lichen Samples (excluding MP_I and MP_II)
#### Calculate and plot pearson's correlation coefficients between Lichen samples
```{r fig.width = 10, fig.height = 10}
lich_counts<- counts_tab[rowSums(counts_tab[c(2:14,16:19)])>0,] #remove rows which have 0 in all samples
lich<-cor(lich_counts[c(2:14,16:19)], method="pearson",use="pairwise.complete.obs")
lich<-cor(counts_tab[c(2:14,16:19)], method="pearson",use="pairwise.complete.obs")
ord_lich<-c("XBA1","XBA2","XSA1","XSA2","XSA2_2","XTA1","XTA2","XBC1","XBC2","XMC2","XSC1",
           "XSC2","XTC2","XBE1","XBE2","XSE2","XTE2")
lich1 <- lich[, ord_lich]
lich1<-lich1[ord_lich, ]

# For correlation plot
corrplot(lich1,method="color",type="full",
         main="", is.corr=F, col=col2(200),addgrid.col = 0,tl.cex = 1, tl.col = "black")


# For multidimensional scaling
plotMDS(lich_counts[c(2:14,16:19)],main="Lichen samples") 

# For heatmap
heat<-lich_counts[c(1:14,16:19)]
heat1<-column_to_rownames(heat, var = "target_id")
exp_matrix<-data.matrix(heat1)
mycols = colorRampPalette(c("blue","white","red"))(100) 
hc_rows <- hclust(as.dist(1-cor(t(exp_matrix), method="pearson")), method="average") # for clustering genes
hc_cols <- hclust(as.dist(1 - cor(exp_matrix, method = "pearson")), method = "average") # to arrange the sample names at x-axis according to hierarchical clustering
htmap<-heatmap.2(exp_matrix, Colv=as.dendrogram(hc_cols),Rowv=as.dendrogram(hc_rows),dendrogram = "both",trace = "none", 
                 col = mycols, keysize = 1, key.title = "Title",
                 scale = c("row"), cexRow = 0.75, cexCol = 1, srtCol = NULL,  
                 srtRow = NULL, adjRow = c(0,NA), sepcolor="white" ,margins = c(7,15), 
                 lwid = c(2,8),
                 lhei= c(2,8), main ="lich samples", na.color="black") #this is the final one for the heat map


```

#### Conclusions:
* Samples group based on substrate (PC1), with barks and twigs grouping together
* Some grouping based on the thallus part is present on PC2, where apothecia are mostly in the upper part of the plot

## 3. Culture Samples
#### Calculate and plot pearson's correlation coefficients between _Xanthoria_ culture samples
```{r fig.width = 10, fig.height = 10}
xan_counts<- counts_tab[rowSums(counts_tab[c(21:32)])>0,] #remove rows which have 0 in xan samples
xan<-cor(xan_counts[c(21:32)], method="pearson",use="pairwise.complete.obs")
ord_xan<-c("KS48XB1","KS48XB2","KS48XB3",
           "KS9XB1","KS9XB2","KS9XB3","S_21XB1","KS21XB1","S_21XB3","S_42XB1","S_42XB2","S_42XB3") #arrange order according to biological replicates (optional)
xan1 <- xan[, ord_xan]
xan1<-xan1[ord_xan, ]
corrplot(xan1,method="color",type="full",is.corr=F, col=col2(200),addgrid.col = 0,tl.cex = 1, tl.col = "black")

# For multidimensional scaling
plotMDS(xan_counts[c(21:32)],main="Xanthoria samples") 

# For heatmap
heat<-xan_counts[c(1,21:32)]
heat1<-column_to_rownames(heat, var = "target_id")
exp_matrix<-data.matrix(heat1)
mycols = colorRampPalette(c("blue","white","red"))(100) 
hc_rows <- hclust(as.dist(1-cor(t(exp_matrix), method="pearson")), method="average") # for clustering genes
hc_cols <- hclust(as.dist(1 - cor(exp_matrix, method = "pearson")), method = "average") # to arrange the sample names at x-axis according to hierarchical clustering
htmap<-heatmap.2(exp_matrix, Colv=as.dendrogram(hc_cols),Rowv=as.dendrogram(hc_rows),dendrogram = "both",trace = "none",  
                 col = mycols, keysize = 1, key.title = "Title",
                 scale = c("row"), cexRow = 0.75, cexCol = 1, srtCol = NULL,  
                 srtRow = NULL, adjRow = c(0,NA), sepcolor="white" ,margins = c(7,15), 
                 lwid = c(2,8),
                 lhei= c(2,8), main ="Xanthoria samples", na.color="black") #this is the final one for the heat map
```

#### Conclusions:
* The 2 dpi samples remain well isolated from the rest
* Samples mostly group together based on the time point, with one exception
* KS21XB1 is a bit different from the rest 21 dpi samples

## 4. ANI between the mycobiont transcriptome and the reference genome


## 5. Next steps:
* Compare all culture samples against all lichen samples (except for MP_I and MP_II)
* For the lichen samples, do pairwise comparisons between apothecia, edge, and center. Include all samples (except for MP_I and MP_II), even from incomplete sets
* For the culture samples, do pairwise comparisons between time points



